---
title: "Ahmed Besbes' "How to score 0.8134 in Titanic Kaggle Challenge" in the R Ecosystem"
author: "Greg van Lunteren"
date: "8/21/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This is meant to be a companion to Ahmed Besbes' wonderful blog post on building a model for Kaggle's Titanic introductory challenge. He uses Python in his post, and this is meant to do approximately the same steps, but in the R ecosystem. Ahmed's post can be found here: http://ahmedbesbes.com/how-to-score-08134-in-titanic-kaggle-challenge.html and the Kaggle competition can be found here: https://www.kaggle.com/c/titanic


# Exploratory Data Analysis

Ahmed breaks this section up into several substeps:
-Data extraction: we'll load the dataset and have a first look at it.
-Cleaning: we'll fill in missing values.
-Plotting: we'll create some interesting charts that'll (hopefully) spot correlations and hidden insights in the data.
-Assumptions: we'll formulate hypotheses from the charts.

###Import relevant libraries:
```{r libraryimports, echo=TRUE, include=FALSE}
install.packages("tidyverse")
library("tidyverse")
```

###Import the training and test data:
```{r dataimports}
setwd("~/Documents/Programming/R/Titanic")
traindata <- read_csv("Titanictrain.csv", col_names = TRUE)
testdata <- read_csv("Titanictest.csv", col_names = TRUE)
```

We want to make sure that read_csv imported our data in the way we would like, so we should review the results:
  PassengerId = col_integer(),
  Survived = col_integer(),
  Pclass = col_integer(),
  Name = col_character(),
  Sex = col_character(),
  Age = col_double(),
  SibSp = col_integer(),
  Parch = col_integer(),
  Ticket = col_character(),
  Fare = col_double(),
  Cabin = col_character(),
  Embarked = col_character()
This all looks good--Now we're ready to take a look inside.

```{r datahead, include=TRUE, echo=TRUE}
head(traindata)
```

There are clearly some missing values, so lets get some numbers on what we're missing:

```{r datahead, include=TRUE, echo=TRUE}
summary(traindata)
```

There are 177 missing age numbers, so we will need to provide values for those. The blog post suggests using the median instead of the mean in this case, so that's what we will do:
```{r fillingblankages include=TRUE, echo=TRUE}
traindata$Age[is.na(traindata$Age)] = median(traindata$Age, na.rm=TRUE)
```
This fills every blank with the median age of 28.

###Graphs

Convert 0/1's to Lived/Died, etc, because ggplot2 doesn't handle 1/0 binaries very well, and make a percentage table
```{r graph_friendly_data, echo=TRUE}
traindataplot <- traindata
traindataplot$Survived[traindataplot$Survived == 1] <- "Lived"
traindataplot$Survived[traindataplot$Survived == 0] <- "Died"
traindataplot$Embarked[traindataplot$Embarked == "S"] <- "England"
traindataplot$Embarked[traindataplot$Embarked == "Q"] <- "Ireland"
traindataplot$Embarked[traindataplot$Embarked == "C"] <- "France"
traindataplot$Pclass[traindataplot$Pclass == "1"] <- "1st"
traindataplot$Pclass[traindataplot$Pclass == "2"] <- "2nd"
traindataplot$Pclass[traindataplot$Pclass == "3"] <- "3rd"

#stats <- traindataplot %>%
#  group_by(Age) %>%
#  summarise(n = )

```




```{r initialGraphs, include=TRUE, echo=TRUE}

SurvSex <- ggplot(data = traindataplot,aes(Sex, fill = Survived)) + geom_bar()
SurvAge <- ggplot(data = traindataplot,aes(Age, fill = Survived)) + geom_histogram(binwidth = 3)
SurvFare <- ggplot(data = traindataplot,aes(Fare, fill = Survived)) + geom_histogram(binwidth = 10)
SurvDepart <- ggplot(data = traindataplot,aes(Embarked, fill = Survived)) + geom_bar()
plot(SurvSex)
plot(SurvAge)
plot(SurvFare)
plot(SurvDepart)
SurvAgeFare <- ggplot(data = traindataplot,aes(Age,Fare,color = Survived)) + geom_point()
plot(SurvAgeFare)
SurvClass <- ggplot(data = traindataplot,aes(Pclass,fill=Survived)) + geom_bar()
plot(SurvClass)
```

[COMMENTARY HERE]

# Feature Engineering
The basic analyses on the data in the previous section yielded some insights, but Ahmed outlines a number of fields that could yield more data with some creative processing. This will include pulling titles out of names, filling in blanks, replacing age more intelligently, and other techniques to get the most from the data.

Before anything else, it's important to combine the training and test datasets so that the datasets match after all the feature engineering steps.

We'll need to add the "Survived" column to the Test dataset in order to merge them. We want to do this in a way where it will be easy to separate them afterwards:
```{r mergeTrainTest,echo=TRUE}
testdata$Survived <- c(1:nrow(testdata))
testdata$Survived <- NA
testdata$Set <- "Test"
traindata$Set <- "Train"
fullset <- rbind(traindata,testdata)
summary(fullset)

```
The combination was successful, as the number of rows (1309) is equal to the sum of the training (891) and test (418) datasets. Note that we still have NA's in a few columns: Fare, Embarked (from the graph section), and Age (which we need to redo according to Ahmed's stratifications)

## Extracting Passenger Titles
Each name has a title in it, from standard titles like Mr, Miss, Mrs, and Dr to more unusual, like Master, Major, Dona, or Jonkheer. The title dictionary that Ahmed created is as follows:

 Title_Dictionary = {
                        "Capt":       "Officer",
                        "Col":        "Officer",
                        "Major":      "Officer",
                        "Jonkheer":   "Royalty",
                        "Don":        "Royalty",
                        "Sir" :       "Royalty",
                        "Dr":         "Officer",
                        "Rev":        "Officer",
                        "the Countess":"Royalty",
                        "Dona":       "Royalty",
                        "Mme":        "Mrs",
                        "Mlle":       "Miss",
                        "Ms":         "Mrs",
                        "Mr" :        "Mr",
                        "Mrs" :       "Mrs",
                        "Miss" :      "Miss",
                        "Master" :    "Master",
                        "Lady" :      "Royalty"
                        }

[[[Obviously this isn't quite in the format that R can use (Tuples), so we can convert to.......]]]




## Adding the Age



